---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# knntuner

<!-- badges: start -->
<!-- badges: end -->

The goal of knntuner is to allow for more efficient cross-validation and tuning for k parameters within KNN problems. Often, these processes require several lines of code in a repetitive manner, and knntuner's goal is to streamline the process.

In the example below, make sure to run each of the code chunks in the same order that they are presented below - they respresent the basic statistical machine learning workflow. Please see the vignette on the package website for more detailed information on each package function, and ways to customize their outputs.

## Installation

You can install the development version of `knntuner` from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
# install.packages("devtools")
devtools::install_github("ADC-405-S26/knntuner")
```

## Example

This is a basic example which shows how `knntuner` can be used in the STATS 205 KNN model building and tuning workflow.

```{r example}
library(knntuner)
library(rsample)
library(caret)
library(recipes)
# necessary packages
```

In this package, there is a small simulated dataset of 100 observations in 6 variables. The data measure student characteristics, and track their performance. A head of the dataset can be seen below, and further information on each variable can be found in the vignette on the package website.

```{r}
head(student_knn_data) # simulated dataset as a part of knntuner (100 observations)
```


First, start with the standard split of data into a training set and a test set. This example uses the student dataset built into the package.

```{r}
set.seed(208)

knntuner_data_split <- initial_split(data = student_knn_data, prop = 0.8) # setting training data to be 80%, and test data to be 20%, and saving these settings as an object

student_train <- training(knntuner_data_split) # training data
student_test  <- testing(knntuner_data_split) # test data
```

Next, create a recipe to remove the ID column, dummy-code categorical predictors, and scale the numeric predictors for KNN, to prepare the data for analysis.

```{r}
student_recipe <- recipe(pass ~ ., data = student_train) |>
  step_rm(id) |> # ID will not be used as a predictor
  step_dummy(all_nominal_predictors()) |> # dummy encoding nominal categorical variables
  step_center(all_numeric_predictors()) |> # normalizing numeric values step 1
  step_scale(all_numeric_predictors()) # normalizing numeric values step 2
```


Then, define your cross-validation specifications with the `cv_control` function.

```{r}
cv <- cv_control() # creating the cv specifications and storing them as cv
```

Next, you can train and cross-validate your model using the specifications that you set. Firstly, use the `kgrid` function to quickly create a dataframe to hold the cross-validation results for each k value. `kgrid` allows you to choose the minimum and maximum k values to test, as well as the step size to take.

```{r}
set.seed(208)

k_grid <- kgrid(1, 9, 2) # knntuner function here, creating the desired kgrid and storing it as an object

student_model <- train(
  student_recipe,
  data = student_train,
  method = "knn",
  trControl = cv, # knntuner object here!
  tuneGrid = k_grid, # knntuner object here!
  metric = "Sens"
)
```

Finally, use the `bestk` function to identify the best performing k value to choose for your final model! Now, you can move on to training your final model, and analyzing its performance on the test data.

```{r}
bestk(student_model) # finding the optimal k value!
```


