---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# knntuner

<!-- badges: start -->
<!-- badges: end -->

The goal of knntuner is to allow for more efficient cross-validation and tuning for k parameters within KNN problems.

## Installation

You can install the development version of `knntuner` from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
# install.packages("devtools")
devtools::install_github("evantoth1/ADC-405-S26", subdir = "knntuner")
```

## Example

This is a basic example which shows how `knntuner` can be used in the STATS 205 KNN model building and tuning workflow.

```{r example}
library(knntuner)
library(rsample)
library(caret)
library(recipes)
# necessary packages
```

First, start with the standard split of data into a training set and a test set. This example uses the student dataset built into the package.

```{r}
set.seed(208)

knntuner_data_split <- initial_split(data = student_knn_data, prop = 0.8)

student_train <- training(knntuner_data_split)
student_test  <- testing(knntuner_data_split)
```

Next, create a recipe to remove the ID column, dummy-code categorical predictors, and scale the numeric predictors for KNN, to prepare the data for analysis.

```{r}
student_recipe <- recipe(pass ~ ., data = student_train) |>
  step_rm(id) |>
  step_dummy(all_nominal_predictors()) |>
  step_center(all_numeric_predictors()) |>
  step_scale(all_numeric_predictors())
```


Then, define your cross-validation specifications with the `cv_control` function.

```{r}
cv <- cv_control() # the default arguments are a "repeatedcv" method, 5 folds, 1 repeat, classProbs = TRUE, and summaryFunction = caret::twoClassSummary
```

Next, you can train and cross-validate your model using the specifications that you set. Firstly, use the `kgrid` function to quickly create a dataframe to hold the cross-validation results for each k value. `kgrid` allows you to choose the minimum and maximum k values to test, as well as the step size to take.

```{r}
set.seed(208)

k_grid <- kgrid(1, 9, 2) # knntuner function here!

student_model <- train(
  student_recipe,
  data = student_train,
  method = "knn",
  trControl = cv,
  tuneGrid = k_grid,
  metric = "Sens"
)
```

Finally, use the `bestk` function to identify the best performing k value to choose for your final model! Now, you can move on to training your final model, and analyzing its performance on the test data.

```{r}
bestk(student_model)
```


