---
title: "Intro to knntuner"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Intro to knntuner}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Intro to knntuner

## Motivation

KNN model workflows often involve repeatedly creating cross-validation controls, tuning grids, and extracting the best-performing k value in tedious repetitions. The knntuner package provides several small helper functions to help streamline this process.

```{r setup}
library(knntuner)
```

## Installation

You can install the development version of `knntuner` from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
# install.packages("devtools")
devtools::install_github("ADC-405-S26/knntuner")
```

Common packages that knntuner works with are loaded below.

```{r}
library(knntuner)
library(rsample)
library(caret)
library(recipes)
```


## Example data

In this package, there is a small simulated dataset of 100 observations in 6 variables. The data measure student characteristics, and tracks their performance. The variables included can be seen below.

* `id` - The ID of the student
* `age` - The age, in years, of the student
* `study_hours` - The average amount of hours per week a student spends studying outside of class
* `score` - The student's test score, out of 100
* `private_tutoring` - Indicates whether the student is privately tutored, outside of school
* `pass` - Indicates whether the student meets the threshold to pass, with a score at or above 70.

```{r}
summary(student_knn_data)
```


## Train/test split

When creating and evaluating a model, it is important to begin by splitting your entire dataset into 2 distinct sections. One section is called the training data, where the model is trained. The other section is called the test data, where the trained model's performance can be evaluated. These two should be kept separate, to prevent the models trained on training data from having inflated performances on the test data.

This is one way to split the data.

```{r}
set.seed(208)

knntuner_data_split <- initial_split(data = student_knn_data, prop = 0.8)

student_train <- training(knntuner_data_split)
student_test  <- testing(knntuner_data_split)
```


## Preprocessing with a recipe

Often when building models, it is important to make small computations, and clean or scale predictor variables for better model performance. This process is called feature engineering. In this example, for the KNN method, because it is a distance-based method, it is important to scale the numeric predictors. This is accomplished in the recipe in the code chunk below, as well as removing the `id` variable from the analysis, because it is not relevant to the analysis.

```{r}
student_recipe <- recipe(pass ~ ., data = student_train) |>
  step_rm(id) |>
  step_dummy(all_nominal_predictors()) |>
  step_center(all_numeric_predictors()) |>
  step_scale(all_numeric_predictors())
```


## Creating cross-validation settings

The first function in `knntuner` comes into play here, called `cv_control`. It allows users to design and store their cross-validation specifications quickly and efficiently. Also, `cv_control` comes with default inputs commonly used in STATS 205. These are listed below.

* `method` - Identifies the type of cross-validation. `"repeatedcv"` is the default
* `number` - Specifies the number of folds to make in the training data during cross-validation. `5` is the default
* `repeats` - Specifies the amount of times to iterate through the cross-validation process. `1` is the default
* `classProbs` - Indicates whether class probabilities should be computed. TRUE is the default.
* `summaryFunction` - Specifies which set of metrics to calculate during cross-validation. `caret::twoClassSummary` is the default.

Thus in some instances, users only need to code in `cv_control()` to create their objects.

```{r}
cv <- cv_control()
```


## Creating a k grid

The second function in `knntuner` allows users to quickly create a k grid to store their cross-validation metrics for each value of k that is tested. Users can specify which k values they want to test using the arguments below.

* `min_k` - Specifies the minimum k value to test
* `max_k` - Specifies the maximum k value to test
* `step` - Specifies the amount to increase k by after each cross-validation process

The example below creates a k grid for k values 1, 3, 5, 7, and 9.

```{r}
k_grid <- kgrid(1, 9, 2) # 1 is the min_k value, 9 is the max_k value, and the step size is 2
k_grid
```


## Training the KNN model

Now, we can actually train the model and apply cross-validation after the preparation. Here, we can use `caret::train()` with the recipe, the `cv` and `k_grid` objects that we created. These act as inputs in the function, and allow for a more efficient application in `caret`. Note that in this case, the metric used to score each model below is the sensitivity.

```{r}
set.seed(208)

student_model <- train(
  student_recipe,
  data = student_train,
  method = "knn",
  trControl = cv, # knntuner object here!
  tuneGrid = k_grid, # knntuner object here!
  metric = "Sens"
)
```


## Extracting the best k

Now that we have applied the cross-validation, we can look through each value of k, and search for the k that performs the best according to how we evaluate it (sensitivity here). The `bestk` function allows us to quickly and easily find and identify the best performing k value, as well as its corresponding metrics.

There is only 1 input:

* `model` - Specifies the model results from which to find the best `k`

```{r}
bestk(student_model)
```

In this example, the k value of 7 performed the best!

## Next steps

Now that we have found an optimal k value, the next steps in the process would be to train the model on all of the training data, using a `k` value of 7. Then we can use that model to make predictions on the test data, and evaluate how accurate the model is overall.


